---
title: "What Makes a Song Award-Winning?"
author: "Cristian Granchelli, Joan Orellana Rios, Cameron Kelahan"
date: "2023-08-25"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Can we use attributes and metrics of songs to predict their likelihood of winning a Grammy?

What common traits, if any, do award winning songs contain? Can we look at intrinsic traits of songs, combined with metrics defined by Spotify, to determine award winning musical features? In this paper, we break down our data collection, data processing, and data analysis of a dataset of roughly 1,000 popular songs, both award-winning and not.

## The Data

### Data Collection and Selection

**JOAN TO DO**: Write how we chose the 1,000 songs originally; how we ended up with 867; and which Grammy winners we stuck with (only Grammy winners, or also Grammy nominees?)

```{r warning = FALSE, echo = TRUE, message = FALSE}
# Set the working directory to this file's folder
library("rstudioapi")
setwd(dirname(getActiveDocumentContext()$path))
load("final_df_n_str.RData")

Sys.setenv(LANG = "en") 

# Load necessary libraries
library(pROC)
library(MASS)
library(ROSE)
library(confintr)
library(ggplot2)
library(correlation)
library(corrplot)
library(class)
library(caret)
library(glmnet)
```

```{r, echo = TRUE, results = 'hide'}
# Selecting the relevant variables
data = final_df_n_str
data = data[,c("track_name", "artist_name", "IsWinner", "Year","year",
               "followers", "acousticness", "danceability", "duration_ms",
               "energy", "instrumentalness", "key", "liveness", "loudness",
               "mode", "tempo", "time_signature", "valence")]

# Merge the two year variable
data$Year[data$Year == "Undefined"] <- data$year[data$Year == "Undefined"]
data = data[,c("track_name","artist_name", "IsWinner", "Year", "followers",
               "acousticness", "danceability", "duration_ms",
               "energy", "instrumentalness", "key", "liveness", "loudness",
               "mode", "tempo", "time_signature", "valence")]

# Eliminating duplicates
data$track_name == "Closing Time"
data$track_name == "Smells Like Teen Spirit"
data$track_name == "Don't Wanna Fight"
data[914, ]
data[789,]
data[669,]

data = data[-c(669, 789, 914),]

sum(data$Year < 1992)
nrow(data)
data = data[!data$Year < 1992,]

# Creating row names

names = paste0(data$track_name, " - ", data$artist_name)

# Eliminating unusable variables
data = data[,c("IsWinner", "Year", "followers", "acousticness",
               "danceability", "duration_ms", "energy",
               "instrumentalness", "key", "liveness", "loudness", "mode",
               "tempo", "time_signature", "valence")]
data = cbind(names = names, data)

# Casting variables
data$IsWinner[data$IsWinner == "Winner"] = 1
data$IsWinner[data$IsWinner == "Nominee"] = 1
data$IsWinner[data$IsWinner == "Nothing"] = 0
data$IsWinner = as.integer(data$IsWinner)
data$Year = as.integer(data$Year)
data$mode = as.factor(data$mode)
data$key = as.factor(data$key)
data$time_signature = as.factor(data$time_signature)

# Giving row names
summary(data)
summary(data$IsWinner)
```

### Explanation of Variables

In order to perform analysis of songs, we decided to use metrics that are intrinsic to music as well as artificial metrics created and measured by the music streaming giant Spotify. The intrinsic metrics we used were: duration, musical key, modality (major or minor key), tempo, time signature, and genre. Spotify also uses what they call ["audio features"](https://developer.spotify.com/documentation/web-api/reference/get-several-audio-features) (in the table below) to perform their own analysis of songs when creating playlists, suggesting music, etc. We used these professionally manufactured metrics to bolster the intrinsic metrics and increase our insight into what might make a song award-winning.

Audio Feature | Definition
------------- | -------------
Acousticness  | A confidence measure from 0.0 to 1.0 of whether the track is acoustic. 1.0 represents high confidence the track is acoustic.
Danceability  | Danceability describes how suitable a track is for dancing based on a combination of musical elements including tempo, rhythm stability, beat strength, and overall regularity. A value of 0.0 is least danceable and 1.0 is most danceable.
Energy        | Energy is a measure from 0.0 to 1.0 and represents a perceptual measure of intensity and activity. Typically, energetic tracks feel fast, loud, and noisy. For example, death metal has high energy, while a Bach prelude scores low on the scale. Perceptual features contributing to this attribute include dynamic range, perceived loudness, timbre, onset rate, and general entropy.
Instrumentalness | Predicts whether a track contains no vocals. "Ooh" and "aah" sounds are treated as instrumental in this context. Rap or spoken word tracks are clearly "vocal". The closer the instrumentalness value is to 1.0, the greater likelihood the track contains no vocal content. Values above 0.5 are intended to represent instrumental tracks, but confidence is higher as the value approaches 1.0.
Liveness      | Detects the presence of an audience in the recording. Higher liveness values represent an increased probability that the track was performed live. A value above 0.8 provides strong likelihood that the track is live.
Loudness      | The overall loudness of a track in decibels (dB). Loudness values are averaged across the entire track and are useful for comparing relative loudness of tracks. Loudness is the quality of a sound that is the primary psychological correlate of physical strength (amplitude). Values typically range between -60 and 0 db.
Speechiness   | Speechiness detects the presence of spoken words in a track. The more exclusively speech-like the recording (e.g. talk show, audio book, poetry), the closer to 1.0 the attribute value. Values above 0.66 describe tracks that are probably made entirely of spoken words. Values between 0.33 and 0.66 describe tracks that may contain both music and speech, either in sections or layered, including such cases as rap music. Values below 0.33 most likely represent music and other non-speech-like tracks.
Valence       | A measure from 0.0 to 1.0 describing the musical positiveness conveyed by a track. Tracks with high valence sound more positive (e.g. happy, cheerful, euphoric), while tracks with low valence sound more negative (e.g. sad, depressed, angry).

Although Spotify does not openly share how they determine these metrics, we found them suitable to assist in our analysis.

As a final processing step, we split the data into training and test datasets. The training dataset contains 80% of the original dataset, and the remaining 20% of the data in the test dataset will be used to test tour model against after we have trained it. It is very important to test the model on never-before-seen data to determine not only how well the model performs, but also how well the model can generalize.

```{r, echo = TRUE, eval = TRUE}
# Splitting training and test set
training_size = floor(0.8 * nrow(data))
set.seed(42)
train_ind = sample(seq_len(nrow(data)), size = training_size)
training_set = data[train_ind,]
test_set = data[-train_ind,]

summary(training_set)

# Checking if the ratio is preserved
sum(data$IsWinner == 1)/ sum(data$IsWinner == 0)
sum(training_set$IsWinner == 1)/ sum(training_set$IsWinner == 0)

training_set
```

## Exploratory Data Analysis

### Relationship Between Independent Variables

At first, we took a look at the continuous variables.

```{r}
attach(training_set)

# Correlations between continuous variables
cor_matrix = cor(training_set[,c(-1, -2, -10, -13, -15)])
corrplot(cor_matrix)
pairs(training_set[,c(-1, -2, -10, -13, -15)], lower.panel = panel.smooth)
```

WHY CAN'T I GET THIS PAIRS() PDF TO INSERT?? Maybe it is inserted, but too big??
```{r}
knitr::include_graphics("yourPlot.pdf", error = FALSE)
```

We looked at the association measure for categorical variables utilizing Cramer's V, which is a normalized version of the chi-square statistic.

(CRISTIAN IS THIS CORRECT? EXPAND?)

```{r, warning=FALSE}
# Association measure for categorical variables (Cramer's V is a normalized 
# version of the chi-square statistics)
cramersv(matrix(c(as.numeric(key), as.numeric(mode)), ncol = 2))
cramersv(matrix(c(as.numeric(key), as.numeric(time_signature)), ncol = 2))
cramersv(matrix(c(as.numeric(mode), as.numeric(time_signature)), ncol = 2))
```

Next, we looked for associations between each of the categorical variables (Key, Mode, and Time Signature) and all of the continuous variables. Some of these were significant, meaning... (ASK CRISTIAN ABOUT THIS)

```{r, warnings = FALSE}
# Association between continuous and categorical variables

# Key
fol_key.aov <- aov(followers ~ key)
summary(fol_key.aov) # SIGNIFICANT

aco_key.aov <- aov(acousticness ~ key)
summary(aco_key.aov)

dan_key.aov <- aov(danceability ~ key)
summary(dan_key.aov) # SIGNIFICANT

dur_key.aov <- aov(duration_ms ~ key)
summary(dur_key.aov)

ene_key.aov <- aov(energy ~ key)
summary(ene_key.aov) # SIGNIFICANT

ins_key.aov <- aov(instrumentalness ~ key)
summary(ins_key.aov)

liv_key.aov <- aov(liveness ~ key)
summary(liv_key.aov)

loud_key.aov <- aov(loudness ~ key)
summary(loud_key.aov)

tem_key.aov <- aov(tempo ~ key)
summary(tem_key.aov)

val_key.aov <- aov(valence ~ key)
summary(val_key.aov)


# Mode
fol_mode.aov <- aov(followers ~ mode)
summary(fol_mode.aov) # SIGNIFICANT

aco_mode.aov <- aov(acousticness ~ mode)
summary(aco_mode.aov) # SIGNIFICANT

dan_mode.aov <- aov(danceability ~ mode)
summary(dan_mode.aov)

dur_mode.aov <- aov(duration_ms ~ mode)
summary(dur_mode.aov)

ene_mode.aov <- aov(energy ~ mode)
summary(ene_mode.aov) # SIGNIFICANT

ins_mode.aov <- aov(instrumentalness ~ mode)
summary(ins_mode.aov)

liv_mode.aov <- aov(liveness ~ mode)
summary(liv_mode.aov)

loud_mode.aov <- aov(loudness ~ mode)
summary(loud_mode.aov) # SIGNIFICANT

tem_mode.aov <- aov(tempo ~ mode)
summary(tem_mode.aov)

val_mode.aov <- aov(valence ~ mode)
summary(val_mode.aov)


# Time signature
fol_time.aov <- aov(followers ~ time_signature)
summary(fol_time.aov)

aco_time.aov <- aov(acousticness ~ time_signature)
summary(aco_time.aov) # SIGNIFICANT

dan_time.aov <- aov(danceability ~ time_signature)
summary(dan_time.aov) # SIGNIFICANT

dur_time.aov <- aov(duration_ms ~ time_signature)
summary(dur_time.aov) # SIGNIFICANT

ene_time.aov <- aov(energy ~ time_signature)
summary(ene_time.aov) # SIGNIFICANT

ins_time.aov <- aov(instrumentalness ~ time_signature)
summary(ins_time.aov) # SIGNIFICANT

liv_time.aov <- aov(liveness ~ time_signature)
summary(liv_time.aov)

loud_time.aov <- aov(loudness ~ time_signature)
summary(loud_time.aov) # SIGNIFICANT

tem_time.aov <- aov(tempo ~ time_signature)
summary(tem_time.aov) # SIGNIFICANT

val_time.aov <- aov(valence ~ time_signature)
summary(val_time.aov) # SIGNIFICANT
```
The fact that x, y, z, etc. were significant leads us to etc. etc. etc. CHAT WITH CRISTIAN ABOUT THIS TOO.

TALK ABOUT PARTIAL CORRELAtIONS
```{r, warnings = FALSE}
# Partial correlations
correlation(training_set[,c(-1, -2, -10, -13, -15)], partial = TRUE)

# Plots of variables with the largest partial correlation
ggplot(data = training_set, aes(danceability, valence)) + geom_jitter(color = "blue")
ggplot(data = training_set, aes(loudness, energy)) + geom_jitter(color = "blue")
ggplot(data = training_set, aes(acousticness, energy)) + geom_jitter(color = "blue")


#Weird song veeeeeeeeeeeeeeeeeeeeeeeery long 
which.max(data$duration_ms)
data[504, ]
```

TALK ABOUT WHY SOME MAY BE SIGNIFICANT; WHAT THAT MEANS.

## Including Plots

You can also embed plots, for example:

```{r pressure, echo=FALSE}
plot(pressure)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
